Inducing Domain Specific Languages for Bayesian Program Learning

### 2018-02-05 paper.tex
introduction rewrite

Imagine an agent faced with a new domain of unfamiliar tasks: think
playing new kinds of videogames, authoring Python code, or navigating
through mazes.  Our agent has at its disposal a basis of primitive
actions it can compose to propose solutions to these tasks, but it is
no idea what kinds of primitives are appropriate for which tasks
nor does it know the higher-level vocabulary in which solutions are
best expressed.  How can our agent learn to solve problems in this new
domain?

The AI and machine learning literature contains two broad takes on this problem.
The first take is that the agent should come up with a better representation of the space of solutions,
for example, by inventing new primitive actions: see \emph{options} in reinforcement learning~\cite{stolle2002learning}, the EC algorithm in program synthesis~\cite{Dechter:2013:BLV:2540128.2540316}, or predicate invention in inductive logic programming~\cite{muggleton2015meta}.
The second take is that the agent should learn a discriminative model mapping problems to a distribution over solutions: for example, policy gradient methods in reinforcement learning or neural models of program synthesis~\cite{devlin2017robustfill,balog2016deepcoder}.
Our contribution is a general algorithm for fusing these two takes on the problem:
we propose jointly inducing a representation language, called a \emph{Domain Specific Language} (DSL),
alongside a bottom-up discriminative model that regresses from problems to solutions.
We evaluate our algorithm on four domains:
building Boolean circuits; symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing problems; and functions on lists.
%Because we want a domain-general
%approach,
We model solutions to tasks as
programs,
which frames our problem as one of program induction.

@@

  \node[latent] at (1,3) (d){$\mathcal{D}$};
  \node[latent] at (2.5,3) (t){$\theta$};
  \node[latent] at (1,1.5) (z){$p$};
  \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
  \node[obs] at (1,-0.5) (x) {$x$};
  \edge {z}{x};
  \edge {d,t}{z};
  \plate {}{(tx)(z)(x)}{$x\in X$};
  \draw [->,dashed] (x.east) to[out = 0,in = -90] node(nn){} (tx.south);
  \draw [->,dashed] (tx.west) -- (z.east);

@@

  \end{tikzpicture}
  \caption{DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives with probabilities $\theta$ (Algorithm~\ref{programGenerativeModel}). We observe tasks $x$, which are program inputs/outputs. A neural network $q(\cdot )$ called the \emph{recognition model} regresses from $x$ to a distribution over programs
    ($\theta^{(x)} = q(x)$). Solid arrows correspond to the top-down generative model. Dashed arrows correspond to the bottom-up recognition model.}\label{graphicalModel}
\end{figure}

Our algorithm is best understood as inference in a hierarchical Bayesian model (Figure~\ref{graphicalModel}: solid lines):
A DSL $\mathcal{D}$
determines a distribution with production probabilities $\theta$,
from which a latent program $p$ is sampled,
which finally determines the agent's observations, $x$.
This makes our work in instance of
 \emph{Bayesian Program
  Learning} (BPL; see~\citep{lake2013one,Dechter:2013:BLV:2540128.2540316,ellis2016sampling,DBLP:conf/icml/LiangJK10}).
In tandem we learn a neural network $q$ that regresses from observations to
programs by proposing a distribution over programs $\theta^{(x)} = q(x)$.
We will show that the interaction of the generative and recognition models
is important for solving these program induction problems.

>>>>

Automatically inducing programs from examples is a long-standing goal
of artificial intelligence.  Recent work in program induction show
that deep networks can be very successful at learning to write certain
kinds of programs (e.g., RobustFill:~\cite{devlin2017robustfill} and
DeepCoder:~\cite{balog2016deepcoder}), as can symbolic search
techniques (e.g., Metagol:~\cite{muggleton2015meta} and
FlashFill:~\cite{gulwani2011automating}).
However, the success of these approaches -- both symbolic and neural --
relies critically upon a hand-engineered \emph{Domain Specific Language} (DSL).
The DSL is an inventory of restricted programming primitives,
crafted by an engineer,
which imparts domain specific knowledge about the space of programs.
To what extent can we dispense with hand-engineered DSLs?

This work proposes \emph{learning} DSLs.
We consider the problem of solving many related program induction tasks,
starting with a Lisp-like representation.
We introduce an algorithm, called the \system,
which creates its own DSL by discovering and then reusing
new useful programming idioms and subroutines.

%% Imagine an agent faced with a new domain of unfamiliar tasks: think
%% playing new kinds of videogames, authoring Python code, or navigating
%% through mazes.  Our agent has at its disposal a basis of primitive
%% actions it can compose to propose solutions to these tasks, but it is
%% no idea what kinds of primitives are appropriate for which tasks
%% nor does it know the higher-level vocabulary in which solutions are
%% best expressed.  How can our agent learn to solve problems in this new
%% domain?

%% The AI and machine learning literature contains two broad takes on this problem.
%% The first take is that the agent should come up with a better representation of the space of solutions,
%% for example, by inventing new primitive actions: see \emph{options} in reinforcement learning~\cite{stolle2002learning}, the EC algorithm in program synthesis~\cite{Dechter:2013:BLV:2540128.2540316}, or predicate invention in inductive logic programming~\cite{muggleton2015meta}.
%% The second take is that the agent should learn a discriminative model mapping problems to a distribution over solutions: for example, policy gradient methods in reinforcement learning or neural models of program synthesis~\cite{devlin2017robustfill,balog2016deepcoder}.
%% Our contribution is a general algorithm for fusing these two takes on the problem:
%% we propose jointly inducing a representation language, called a \emph{Domain Specific Language} (DSL),
%% alongside a bottom-up discriminative model that regresses from problems to solutions.
%% We evaluate our algorithm on four domains:
%% building Boolean circuits; symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing problems; and functions on lists.
%% %Because we want a domain-general
%% %approach,
%% We model solutions to tasks as
%% programs,
%% which frames our problem as one of program induction.

@@

  \node[latent] at (0.5,3) (d){$\mathcal{D}$};
  \node[latent] at (1.5,3) (t){$\theta$};
  \node[latent] at (1,1.75) (z){$p$};
%  \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
  \node[obs] at (1,0) (x) {$x$};
  \edge {z}{x};
  \edge {d,t}{z};
  \plate {}{(z)(x)}{$x\in X$};
  \node[align = center] at ([yshift = -1.3cm]x.south) {(a) \\generative \\model};

  \begin{scope}[shift = {(-1,0)}]  
  \node[latent] at (3.5,3) (dx){$\mathcal{D}$};
  \node[latent] at (3.5,1.75) (zp){$p$};
  \node[latent] at (5.5,1.75) (tx){$\theta^{(x)}$};
  \node[obs] at (3.5,0) (xp) {$x$};
  \plate {}{(zp)(xp)(tx)}{$x\in X$};
  \node[align = center] at ([yshift = -1.2cm,xshift = 1cm]xp.south) {(b) \\recognition model};
  \draw [->,red] (xp.east) to[out = 0,in = -90] node(nn){} (tx.south);
  \draw [->,red] (tx.west) -- (zp.east);
  \draw [->,red] (dx.south) -- (zp.north);

@@

  \end{scope}

  \end{tikzpicture}
  \caption{(a) DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives with probabilities $\theta$ (Algorithm~\ref{programGenerativeModel}). Observe tasks $x$ (program inputs/outputs). (b)  neural network $q(\cdot )$, the \emph{recognition model}, regresses from $x$ to a distribution over programs
    ($\theta^{(x)} = q(x)$). Black arrows correspond to the top-down generative model. Red arrows correspond to the bottom-up recognition model.}\label{graphicalModel}
\end{figure}

Our algorithm is best understood as inference in a hierarchical Bayesian model (Figure~\ref{graphicalModel}(a): black lines):
A DSL $\mathcal{D}$, equipped with a real-valued weight vector $\theta$,
determine a distribution over programs
from which a latent program is sampled,
which finally determines the agent's observations.
Our goal will be both to infer the programs that best explain our observations,
and at the same time induce a DSL
that distills the commonalities across all of the programs that solve the tasks.
Inducing the DSL means learning the symbolic structure of a latent programming language,
and so we are confronted with two distinct structure learning problems:
the structure of the programs and the structure of the DSL.

Inference in the generative model in Figure~\ref{graphicalModel}(a) is
difficult because of the vastness of the combinatorial space of all
programs. To make inference tractable our model learns a bottom-up
\emph{recognition model} alongside the top-down generative model
(Figure~\ref{graphicalModel}b).  The recognition model $q$ is a neural
network that regresses from observations to a distribution over
programs.  The neural network $q$ implements an amortized
inference/compiled inference
scheme~\cite{le2016inference,ritchie2016deep} for the generative
model.

How should we perform inference for the model
in~\ref{graphicalModel}(a)?  An intuitive strategy is to exploit the
conditional independence of the programs conditioned on the DSL, and
so to alternate between estimating the DSL ($\mathcal{D},\theta$) and
estimating a posterior distribution over the programs by searching for
programs written using the current DSL.  In practice searching the
space of programs is difficult, so we use the recognition model to
learn how to efficiently search for programs in the current DSL.

This makes our work in instance of
 \emph{Bayesian Program
  Learning} (BPL; see~\citep{lake2013one,Dechter:2013:BLV:2540128.2540316,ellis2016sampling,DBLP:conf/icml/LiangJK10}).
 Alongside this generative model, our algorithm learns a bottom-up recognition model (Figure~\ref{graphicalModel}(b)).
 The recognition model is a neural network $q$ that regresses from programming tasks $x$
 to a distribution over programs.
